Cover letter for the original IJRR submission

### 2017-05-03 body/intro.tex
Tentative complete paper.

In the specific (and simple) case of gaited biped locomotion on flat ground, choosing the effector with which to create a contact is trivial because walking follows a cyclic pattern (left foot always follows right foot).
Efficient tools such as the capture point~\citep{Pratt2006} can be used to compute the next contact location. 
%
The general case of planning contact sequences is more challenging.
At any given time a contact results of a combinatorial discrete choice for both the effector and contact surface, and a continuous choice for the contact location.
Any choice of contact then kinematically and dynamically constrains the possible motion of the robot.
Furthermore there is no analytical way to verify whether this choice brings the robot one step closer to the desired goal or to a dead end, especially in the presence of obstacles.
From this property, we say that the contact manifold is foliated~\citep{simeon-manipulation-04}. 
The foliation prevents the use of efficient sampling-based planners for two reasons.
(i) First, each submanifold of the foliation has a zero measure and cannot be directly sample.
A sample is rather obtained by sampling a ``free flying'' configuration and explicitly projecting it in contact (which is a costly numerical operation).
(ii) Second, the foliated topology turns the exploration by spreading a graph of configurations (probability roadmap, rapidely exploring random trees) into an unefficient random process, where many useless nodes are sampled on parallel submanifolds.
The total algorithm complexity of classical contact planners comes both from the number of nodes of the graph that are sampled during the exploration (ii) and to the cost of the projection when sampling a new configuration (i).
%
For this reason previous contributions having demonstrated acyclic contact locomotion on a real robot are too computationally expensive~\citep{Bretl:2006:MPM:1124573.1124585}. As a result at the DARPA Robotics Challenge, the participants stated that the robots did not use multi contact strategies (except for egress): they relied on unsafe bipedal walking to climb stairs, instead of using the provided handrails to facilitate the motion~\citep{atkensondarpa}.

%Obtaining reasonable computation times is critical for a contact planner to be of any practical use. 
Our work aims at breaking the complexity of the acyclic contact planning problem.  
To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold (i), and the combinatorics of the contact selection problem (ii). 
To that end, we introduce a reduction property: the \textit{reachability condition}.
This condition relaxes the strict contact condition.
It induces a low-dimensional space, called the \textit{contact reachable} space, that has Cartesian topology, i.e. it can be efficiently sampled.
Under some hypothesis, a path in the \textit{contact reachable} space can be easily mapped to the (high-dimensionnal) contact manifold. 
This approach enables us to separate the contact planning problem in two tractable subproblems.
This decoupling presents pros and cons discussed in previous related literature, summarized in the following.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

@@

Additionally to robotics, acyclic motion planning is also a problem of interest in neurosciences, biomechanics, and virtual character animation.
Early contributions in the latter field rely on local adaptation of motion graphs \citep{citeulike:220163}, or ad-hoc construction of locomotion controllers \citep{Pettre:2003:LPD:846276.846313}. These approaches are by definition not able to discover complex behaviors in unforeseen contexts.

The issue of planning acyclic contacts was first completely described by Bretl~\cite{Bretl:2006:MPM:1124573.1124585}. The issue requires the simultaneous handling of two sub-problems, $\mathcal{P}_1$: planning a guide path for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: planning a discrete sequence of equilibrium configurations along the path. A third nontrivial problem, $\mathcal{P}_3$, 
%~ not addressed in this work, 
then consists in interpolating a complete motion between two postures of the contact sequence.  A key issue is to avoid combinatorial explosion when considering at the same time the possible contacts and the potential paths. Bretl's seminal paper proposes a first effective algorithm, able to handle simple situations (such as climbing scenarios), but not applicable to arbitrary environments. Following it, seve\-ral papers have applied this approach in specific situations, limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

Most of the papers that followed the work of Bretl have explored alternative formulations to handle the combinatorics. Two main directions have been explored. \textbf{On the one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial of the complete problem for a differential complexity, at the cost of local convergence \citep{1631739}. A complete example of the potential offered by such approaches was proposed by \cite{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To get reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the method is far from real-time  (about 1 minute of computation for 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. Deits and Tedrake proposed to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered, and equilibrium is not considered~\cite{DBLP:conf/humanoids/DeitsT14}. 
Dai et al.~\cite{dai2014whole} extended the work of Posa et al.~\cite{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for complex interactions.
In addition to the limits of the current implementations, optimization-based approaches only converge locally.

For completeness, we lastly mention a new kind of approach, recently proposed in the computer graphics field \citep{hamalainen_cpbp_2015}. The authors
used black-box physics simulators to perform Model Predictive Control for a humanoid character, and managed to obtain dynamically-consistent motions
at \gls{interactive} frame rates\footnote{We say that a method is \underline{interactive} when the computation time for planning is less than the
execution time.}. While this new approach provides an exciting direction of research, currently the resulting motions
 do not seem applicable to real robots.

@@

Our solution belongs to the class of decoupled approaches, because we believe it is the most promising direction~\citep{DBLP:conf/iser/EscandeKMG08} to break the complexity of acyclic contact planning. Compared to previous approaches, our solution has two main novelties:

\noindent \textbf{Regarding $\mathcal{P}_1$}. We propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample directly the contact manifold, but an approximation of the \textit{contact reachable} space. The \textit{contact reachable} space is a low-dimensional space for which there exists a mapping to the contact manifold.

\noindent \textbf{Regarding $\mathcal{P}_2$},  we propose a fast method to extend a \textit{contact reachable} path into a sequence of whole-body configurations in static equilibrium. This  requires the explicit computation of contact configurations. It is guided by dedicated heuristics that quickly synthesize feasible configurations.

This sequential approach is the key to the efficiency of our method, though it
can result in failures. However, we demonstrate empirically its interest: the high success rate and low computation times allow us to plan (and re-plan upon failure) multi-contact sequences at \gls{interactive} rates.

To further demonstrate the validity of our approach, we show that the generated contact plans  can be successfully executed (problem  $\mathcal{P}_3$), either in simulation or on the real HRP-2 robot. For HRP-2, we detail the complete computation times to address sequentially the three problems, and compare them to related work, demonstrating that our method is orders of magnitude faster.

Finally, we provide an extensive discussion on the consequences of our approach in terms of efficiency and completeness regarding the contact planning problem.

This paper extends our previous conference paper~\citep{tonneauisrr15} in several ways. Our approach is now demonstrated on actual robots and in simulation, contrary to virtual avatars before, in a large variety of contexts. To achieve this we introduce a new criterion to efficiently compute the equilibrium of the robot while accounting for uncertainties of the real world. We provide the complete pseudo code of algorithm. We give the complete procedure to use our planner with any legged robot and provide our source code, allowing any user to directly use our method. We also propose a solution to $\mathcal{P}_3$, making our framework complete. Finally we provide an in-depth analysis of the results we obtain, and a theoretical discussion on the capabilities of the planner and its current limitations.

>>>>

In the specific (and simple) case of gaited biped locomotion on flat ground, choosing the effector with which to create a contact is trivial because walking follows a cyclic pattern (left foot always follows right foot).
Efficient tools such as the capture point~\citep{Pratt2006} can be used to compute the next contact location. 
%
The general case of planning contact sequences is more challenging.
At any given time a contact results of a combinatorial discrete choice for both the effector and contact surface, and a continuous choice for the contact location.
Any choice of contact then kinematically and dynamically constrains the possible motion of the robot.
Furthermore there is no analytical way to verify whether this choice brings the robot one step closer to the desired goal or to a dead end, especially in the presence of obstacles.
From this property, we say that the contact manifold is foliated~\citep{simeon-manipulation-04}. 
The foliation prevents the use of efficient sampling-based planners for two reasons.
(i) First, each submanifold of the foliation has a zero measure and cannot be directly sample.
A sample is rather obtained by sampling a ``free flying'' configuration and explicitly projecting it in contact (which is a costly numerical operation).
(ii) Second, the foliated topology turns the exploration by spreading a graph of configurations (probability roadmap, rapidely exploring random trees) into an unefficient random process, where many useless nodes are sampled on parallel submanifolds.
The total algorithmic complexity of classical contact planners comes from both the number of graph nodes sampled during exploration (ii) and the cost of the projection when sampling new configurations (i).
%
For this reason previous contributions having demonstrated acyclic contact locomotion on a real robot are too computationally expensive~\citep{Bretl:2006:MPM:1124573.1124585}. As a result at the DARPA Robotics Challenge, the participants stated that the robots did not use multi contact strategies (except for egress): they relied on unsafe bipedal walking to climb stairs, instead of using the provided handrails to facilitate the motion~\citep{atkensondarpa}.

%Obtaining reasonable computation times is critical for a contact planner to be of any practical use. 
Our work aims at breaking the complexity of the acyclic contact planning problem.  
We deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold (i), and the combinatorics of the contact selection problem (ii). 
To that end, we introduce a reduction property: the \textit{reachability condition}.
This condition relaxes the strict contact condition.
It induces a low-dimensional space, called the \textit{contact reachable} space, that has Cartesian topology, i.e. it can be efficiently sampled.
Under some hypothesis, a path in the \textit{contact reachable} space can be easily mapped to the (high-dimensionnal) contact manifold. 
This approach enables us to separate the contact planning in two tractable subproblems.
This decoupling presents pros and cons discussed in previous related literature, summarized in the following.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

Additionally to robotics, acyclic contact planning is a problem of interest in neurosciences, biomechanics and computer animation.
Early contributions in the latter field rely on local adaptation of motion graphs \citep{citeulike:220163} or ad-hoc construction of locomotion controllers \citep{Pettre:2003:LPD:846276.846313}. These approaches are by definition not able to discover complex behaviors in unforeseen contexts.

The issue of planning acyclic contacts was first completely described by Bretl~\cite{Bretl:2006:MPM:1124573.1124585}. It requires the simultaneous handling of two sub-problems, $\mathcal{P}_1$: planning a guide path for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: planning a discrete sequence of equilibrium configurations along the path. A third nontrivial problem, $\mathcal{P}_3$, 
%~ not addressed in this work, 
then consists in interpolating a complete motion between two configurations of the contact sequence.  A key issue is to avoid combinatorial explosion when considering at the same time the possible contacts and the potential paths. Bretl's seminal paper proposes a first effective algorithm, able to handle simple situations, but not applicable to arbitrary environments. Following it, seve\-ral papers have applied this approach in specific situations, limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

Most of the papers that followed the work of Bretl have explored alternative formulations to handle the combinatorics. Two main directions have been explored. \textbf{On the one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial of the complete problem for a differential complexity, at the cost of local convergence \citep{1631739}. A complete example of the potential offered by such approaches was proposed by \cite{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To get reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the method is far from real-time  (about 1 minute of computation for 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. Deits and Tedrake proposed to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered, and equilibrium is not considered~\cite{DBLP:conf/humanoids/DeitsT14}. 
Dai et al.~\cite{dai2014whole} extended the work of Posa et al.~\cite{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for complex interactions.
Model predictive control on short horizon of the whole contact dynamcis~\cite{tassa_icra14,hamalainen_cpbp_2015} leads to dynamically-consistent motions at \gls{interactive} frame rates\footnote{We say that a method is \underline{interactive} when the computation time for planning is less than the execution time.}, although with very few look-ahead capabilities and yet not realistically applicable to physical robots.
In addition to the limits of the current implementations, optimization-based approaches only converge locally.

%% For completeness, we lastly mention a new kind of approach, recently proposed in the computer graphics field \citep{hamalainen_cpbp_2015}. The authors
%% used black-box physics simulators to perform Model Predictive Control for a humanoid character, and managed to obtain dynamically-consistent motions
%% at \gls{interactive} frame rates\footnote{We say that a method is \underline{interactive} when the computation time for planning is less than the
%% execution time.}. While this new approach provides an exciting direction of research, currently the resulting motions
%%  do not seem applicable to real robots.

@@

@@

Our solution belongs to the class of decoupled approaches, i.e. we propose specific algorithms to efficiently solve both $\mathcal{P}_1$ and $\mathcal{P}_2$, while relying on state-of-the-art solution to $\mathcal{P}_3$ to obtain the whole movement.
Our main contribution is the definition of a reduction property, the \textit{reachibility condition}.%,  that leads to a large decrease of the overall algorithmic complexity.
Compared to previous approaches, our solution has two main novelties:

\noindent \textbf{Regarding $\mathcal{P}_1$}, we propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample directly the contact manifold, but an approximation of the \textit{contact reachable} space. The \textit{contact reachable} space is a low-dimensional space and can be explored using standard sample-based planning.
Under some conditions, a mapping exists from a trajectory in the \textit{contact reachable} space to the contact manifold.

\noindent \textbf{Regarding $\mathcal{P}_2$}, we propose a fast method to extend a \textit{contact reachable} path into a sequence of whole-body configurations in static equilibrium. 
Thanks to the reachibility condition, the sequence can be extended exactly on the trajectory computed in $\mathcal{P}_1$ (and not along it, like in previous works~\cite{DBLP:conf/iser/EscandeKMG08,Bouyarmane2009}), which makes $\mathcal{P}_2$ a simple 1-D algorithm.
It requires the explicit computation of contact configurations, guided by dedicated heuristics that quickly synthesize feasible configurations.

The reachibility condition is the key to the strict separation between $\mathcal{P}_1$ and $\mathcal{P}_2$, hence to the low complexity of our planner.
It  may also result in failure (i.e. some guide paths $\mathcal{P}_1$ are not extendable in a full sequence $\mathcal{P}_2$).
However, we empirically show by a large number of trials on various scenarios of practical interest that the rate of success is very high, while the algorithm efficiency allows the planner to iterate in case of failure.
All scenarios are validated by planning a whole dynamic trajectory using a state-of-the-art implementation of $\mathcal{P}_3$ which is executed either in simulation or on the real robot.

The paper extends our previous papers presenting the reachibility condition~\cite{tonneauisrr15} and the connection of our contact planner with a solver~$\mathcal{P}_3$~\cite{Carpentier2016}.
In this paper, we introduce a new efficient criterion of the robot balance, provide the complete pseudo-code an release our source code that can be used to directly reproduce our results.
We also provide a solution to $\mathcal{P}_3$.
The planner is validated on robot models of HyQ and HRP-2 (while \cite{tonneauisrr15} used unrealistic avatar models) with complete performance analysis.

%This sequential approach is the key to the efficiency of our method, though it can result in failures. 
%
%However, we demonstrate empirically its interest: the high success rate and low computation times allow us to plan (and re-plan upon failure) multi-contact sequences at \gls{interactive} rates.

%% To further demonstrate the validity of our approach, we show that the generated contact plans  can be successfully executed (problem  $\mathcal{P}_3$), either in simulation or on the real HRP-2 robot. For HRP-2, we detail the complete computation times to address sequentially the three problems, and compare them to related work, demonstrating that our method is orders of magnitude faster.

%% Finally, we provide an extensive discussion on the consequences of our approach in terms of efficiency and completeness regarding the contact planning problem.

%% This paper extends our previous conference paper~\citep{tonneauisrr15} in several ways. Our approach is now demonstrated on actual robots and in simulation, contrary to virtual avatars before, in a large variety of contexts. To achieve this we introduce a new criterion to efficiently compute the equilibrium of the robot while accounting for uncertainties of the real world. We provide the complete pseudo code of algorithm. We give the complete procedure to use our planner with any legged robot and provide our source code, allowing any user to directly use our method. We also propose a solution to $\mathcal{P}_3$, making our framework complete. Finally we provide an in-depth analysis of the results we obtain, and a theoretical discussion on the capabilities of the planner and its current limitations.

### 2017-05-03 body/overview.tex
Tentative complete paper.

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed sequentially: From a given problem (a) we first plan a root guide path (b), before
extending it into a sequence of static equilibrium configurations(c). In the case where step c fails,
our framework invalidates the computed guide path, and restarts the planning from b.

>>>>

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed sequentially: from a given problem (a) we first plan a root guide path (b), before
extending it into a sequence of static equilibrium configurations (c). In the case where step (c) fails,
our framework invalidates the computed guide path, and restarts the planning from (b).

### 2017-05-03 body/rbprm.tex
Tentative complete paper.

% !TEX root =  ../main.tex
%~ \section{Computing a guide path in $C^0_{reach}$ ($\mathcal{P}_1$) }
\section{Root path planning in the contact reachable space}
\label{rbprm}

During the root path planning we only consider the root configuration $\mathbf{q}^0$ defined in the previous Section,
as well as the environment $O$.

Given a start and a goal configurations, we aim at computing a guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow$ $\mathbb{R}^r$ verifying:
\begin{equation*} \label{eq:path}
\forall t \in [0,1], \mathbf{q}^0(t)  \in C_{Equil}^0
\end{equation*}
This means that any root configuration must be extended into a whole-body, static equilibrium configuration.
$C_{Equil}^0$  cannot be described analytically.

The strong hypothesis of this work is that for a large variety of locomotion tasks, we can define a space  $C_{Reach}^0 \simeq C_{Contact}^0$, such that 
\begin{equation} \label{eq:creach}
\forall t \in [0,1], \mathbf{q}^0(t) \in C_{Reach}^0 \Rightarrow \mathbf{q}^0(t)  \in C_{Equil}^0
\end{equation}
% for \gls{cluttered} problems.
We call  $C_{Reach}^0$ the \textit{contact reachable workspace}, and detail its construction in the following.
The validity of this hypothesis is discussed in depth in Section~\ref{sec:discussion}.
%~ We describe the construction of $C_{Reach}^0$ in the following.

@@

\textbf{necessary conditions:}
For a contact to be possible, an obstacle $O_i \subset O$ necessarily intersects the reachable workspace $W(\mathbf{q}^{0})$ of the robot (Figure~\ref{fig:contact_gen}--1). Also the torso of the robot $W^0(\mathbf{q}^{0})$ must necessarily be collision-free.
Therefore we can define an outer approximation  $ \mathcal{C}^0_{\textrm{\it Nec}} \supset$ $C_{Contact}^0$ as: 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Nec}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \textrm{ \textbf{and}}\ W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%~ The inclusion $C_{Contact}^0$ $\subset \mathcal{C}^0_{\textrm{\it Nec}}$ is straightforward, but it is very important.
%~ Planning in $\mathcal{C}^0_{\textrm{\it Nec}}$ allows a strong reduction of the search space, while not discarding 
%~ any root configuration that could lead to a solution.

@@

\textbf{sufficient condition:}
Similarly we can define an inner approximation \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ $C_{Contact}^0$} by considering a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces. 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Suf}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \textrm{ \textbf{and}}\ B^{\textrm{\it Suf}}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}

@@

\subsection{The compromising reachability condition}
\label{sec:scaling}
The ideal shape $B^*, W^0 \subset B* \subset B^{\textrm{\it Suf}}$ that defines a necessary \textbf{and} sufficient condition for contact creation has no explicit definition.
Therefore, we approximate $B^*$ to define the contact reachable space $C_{Reach}^0$.

%~ However, using a shape between $W^0$ and $B^{\textrm{\it Suf}}$ leads to a trade-off between a necessary and a sufficient condition. 
We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
 \label{eq:reachc}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \textrm{ \textbf{and}}\ W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
The parametrization of $s$ defines a trade-off:
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $\mathcal{C}^0_{\textrm{\it Nec}}$.
 %~ We thus consider that $s \geq 1$, since smaller values would only worsen the approximation.
By increasing $s$, the condition can become sufficient, but less and less necessary.  
Eq.~\ref{eq:reachc} thus defines the \textit{reachability condition}. We fix a value $s^*$ for $s$ and define  $C_{Reach}^0 = C^0_{s^*}$.
The computation of $s^*$ is detailed in Section~\ref{sec:params}. 
In Appendix~\ref{app:rom}, we give a generic method to compute the $W$ volumes appearing in the definition of $C_{Reach}^0$.

\subsection{Computing the guide path in $C_{reach}^0$}
$C_{Reach}^0$ can be sampled efficiently thanks to Eq.~\ref{eq:reachc}, and can thus be used with a standard motion planner.
The only significant change is to replace the collision checking with the \textit{reachability condition}.
Our current implementation of these modifications is based on the Bi-RRT planner \citep{770022} provided by the HPP software~\citep{7759083}. \\

@@

This Section has presented a guide path planner for the geometric root of a robot, implemented as a low-dimensional sampling-based 
algorithm. Given start and goal configurations, it outputs a continuous path for the robot's root. 
%~ \deladp{Thanks to the \textit{reachability condition}, we assume that for any configuration in the path there exists a joint configuration that results in static equilibrium.}
%~ Thanks to these modifications, the problem of planning a \gls{contact reachable} path is reduced to a geometric collision-checking problem, of low dimension (6 for HyQ, 8 for HRP-2 that has 2 joints in the torso).
%~ By doing this we can solve the problem with a sampling-based approach, in \gls{interactive} computation times.

>>>>

% !TEX root =  ../main.tex
%~ \section{Computing a guide path in $C^0_{reach}$ ($\mathcal{P}_1$) }
\section{Root path planning in the contact reachable space ($\mathcal{P}_1$)}
\label{rbprm}

During the root path planning we only consider the root configuration $\mathbf{q}^0$ defined in the previous Section,
as well as the environment $O$.
%
Given a start and a goal configurations, we aim at computing a guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow$ $\mathbb{R}^r$ verifying:
\begin{equation*} \label{eq:path}
\forall t \in [0,1], \mathbf{q}^0(t)  \in C_{Equil}^0
\end{equation*}
This means that any root configuration must be extended into a whole-body, static equilibrium configuration.
$C_{Equil}^0$  cannot be described analytically.
%
The strong hypothesis of this work is that for a large variety of locomotion tasks, we can define a space  $C_{Reach}^0 \simeq C_{Contact}^0$, such that 
\begin{equation} \label{eq:creach}
\forall t \in [0,1], \mathbf{q}^0(t) \in C_{Reach}^0 \Rightarrow \mathbf{q}^0(t)  \in C_{Equil}^0
\end{equation}
% for \gls{cluttered} problems.
We call  $C_{Reach}^0$ the \textit{contact reachable workspace}, and detail its construction in the following.
The validity of this hypothesis is discussed in depth in Section~\ref{sec:discussion}.
%~ We describe the construction of $C_{Reach}^0$ in the following.

@@

\noindent\textbf{Necessary conditions:}
For a contact to be possible, an obstacle $O_i \subset O$ necessarily intersects the reachable workspace $W(\mathbf{q}^{0})$ of the robot (Figure~\ref{fig:contact_gen}--1). Also the torso of the robot $W^0(\mathbf{q}^{0})$ must necessarily be collision-free.
Therefore we can define an outer approximation  $ \mathcal{C}^0_{\textrm{\it Nec}} \supset$ $C_{Contact}^0$ as: 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Nec}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \textrm{ \textbf{and}}\ W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%~ The inclusion $C_{Contact}^0$ $\subset \mathcal{C}^0_{\textrm{\it Nec}}$ is straightforward, but it is very important.
%~ Planning in $\mathcal{C}^0_{\textrm{\it Nec}}$ allows a strong reduction of the search space, while not discarding 
%~ any root configuration that could lead to a solution.

@@

\noindent\textbf{Sufficient condition:}
Similarly we can define an inner approximation \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ $C_{Contact}^0$} by considering a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces. 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Suf}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \textrm{ \textbf{and}}\ B^{\textrm{\it Suf}}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}

@@

\subsection{The compromising reachability condition}
\label{sec:scaling}
The ideal shape $B^*, W^0 \subset B* \subset B^{\textrm{\it Suf}}$ that defines a necessary \textbf{and} sufficient condition for contact creation has no explicit definition.
Therefore, we approximate $B^*$ to define the contact reachable space $C_{Reach}^0$.
%
%~ However, using a shape between $W^0$ and $B^{\textrm{\it Suf}}$ leads to a trade-off between a necessary and a sufficient condition. 
We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
 \label{eq:reachc}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \textrm{ \textbf{and}}\ W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
The parametrization of $s$ defines a trade-off:
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $\mathcal{C}^0_{\textrm{\it Nec}}$.
 %~ We thus consider that $s \geq 1$, since smaller values would only worsen the approximation.
By increasing $s$, the condition can become sufficient, but less and less necessary.  
Eq.~\eqref{eq:reachc} thus defines the \textit{reachability condition}. We fix a value $s^*$ for $s$ and define  $C_{Reach}^0 = C^0_{s^*}$.
The computation of $s^*$ is detailed in Section~\ref{sec:params}. 
In Appendix~\ref{app:rom}, we give a generic method to compute the $W$ volumes appearing in the definition of $C_{Reach}^0$.

\subsection{Computing the guide path in $C_{reach}^0$}
$C_{Reach}^0$ can be sampled efficiently thanks to \eqref{eq:reachc}, and can thus be used with a standard motion planner.
The only significant change is to replace the collision checking with the \textit{reachability condition}.
Our current implementation of these modifications is based on the Bi-RRT planner \citep{770022} provided by the HPP software~\citep{7759083}. \\

@@

%This Section has presented a guide path planner for the geometric root of a robot, implemented as a low-dimensional sampling-based 
%algorithm. Given start and goal configurations, it outputs a continuous path for the root of the robot.

The result of this section is a guide path for the root of the robot (problem $\mathcal{P}_1$), computed with simple modification over a bi-RRT planner using the reachability condition.
We now explain how this path of the root can be extended to a sequence of whole-body contact configurations.

%~ \deladp{Thanks to the \textit{reachability condition}, we assume that for any configuration in the path there exists a joint configuration that results in static equilibrium.}
%~ Thanks to these modifications, the problem of planning a \gls{contact reachable} path is reduced to a geometric collision-checking problem, of low dimension (6 for HyQ, 8 for HRP-2 that has 2 joints in the torso).
%~ By doing this we can solve the problem with a sampling-based approach, in \gls{interactive} computation times.
